Red Wine Quality Exploratory Data Analysis by Lam Raga A. Markely
========================================================


```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(corrplot)
library(psych)
library(ggplot2)
library(dplyr)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
wine = read.csv('wineQualityReds.csv')
```

This dataset contains information about various red wine samples with their 
chemical properties and quality. The goal of the analysis is to identify which 
chemical properties influence the quality of red wines.

# Univariate Plots Section

```{r echo=FALSE, Univariate_Plots}
dim(wine)
str(wine)
summary(wine)
```

The red wine dataset consists of 12 features, including 11 input variables 
(fixed.acidity to alcohol), 1 output variable (quality), and 1599 samples. 
The feature X is the index of the dataset. 
The definition of the variables are described in detail 
[here](https://s3.amazonaws.com/udacity-hosted-downloads/ud651/wineQualityInfo.txt).
In the following analysis, the variable fixed acidity will be referred as
tartaric acid, volatile acidity as acetic acid, chlorides as sodium chloride,
and sulphates as potassium sulphate as described in the above link. 

```{r echo=FALSE}
ggplot(wine,aes(quality)) +
  geom_histogram(binwidth=1,color='white') +
  scale_x_continuous(breaks=seq(3,8,1)) +
  xlab('Quality') +
  ylab('Count')

summary(wine$quality)
```

Most of the wine samples have quality 5 and 6. 
The summary statistics are provided above.

```{r echo=FALSE}
ggplot(wine,aes(fixed.acidity)) +
  geom_histogram(color='white',binwidth=1) +
  scale_x_continuous(breaks=seq(4,16,1)) +
  xlab('Tartaric Acid (g/L)') +
  ylab('Count')

summary(wine$fixed.acidity)

```

The tartaric acid concentration distribution is slightly skewed to the right.
The summary statistics are provided above.

```{r echo=FALSE}
ggplot(wine,aes(volatile.acidity)) +
  geom_histogram(color='white',binwidth=0.1) +
  scale_x_continuous(breaks=seq(0.1,1.6,0.1)) +
  xlab('Acetic Acid (g/L)') +
  ylab('Count')

summary(wine$volatile.acidity)
```

The acetic acid concentration distribution is slightly skewed to the right.
The summary statistics are provided above.

```{r echo = FALSE}
ggplot(wine,aes(citric.acid)) +
  geom_histogram(color='white',binwidth=0.05) +
  scale_x_continuous(breaks=seq(0,1.0,0.1)) +
  xlab('Citric Acid (g/L)') +
  ylab('Count')

summary(wine$citric.acid)
```

The citric acid concentration distribution is very skewed to the right.
The summary statistics are provided above.

```{r echo=FALSE}
ggplot(wine,aes(residual.sugar)) +
  geom_histogram(color='white',binwidth=0.5) +
  scale_x_continuous(breaks=seq(0,16,1)) +
  xlab('Residual Sugar (g/L)') +
  ylab('Count')

summary(wine$residual.sugar)
```

The residual sugar concentration distribution is skewed to the right.
The summary statistics are provided above.

```{r echo=FALSE}
ggplot(wine,aes(chlorides)) +
  geom_histogram(color='white',binwidth=0.02) +
  scale_x_continuous(breaks=seq(0,0.7,0.05)) +
  xlab('Sodium Chloride (g/L)') +
  ylab('Count')

summary(wine$chlorides)
```

The sodium chloride concentration distribution is skewed to the right.
The summary statistics are provided above.

```{r echo=FALSE}
ggplot(wine,aes(free.sulfur.dioxide)) +
  geom_histogram(color='white',binwidth=2) +
  scale_x_continuous(breaks=seq(0,80,5)) +
  xlab('Free Sulfur Dioxide (mg/L)') +
  ylab('Count')

summary(wine$free.sulfur.dioxide)
```

The wine sodium chloride concentration distribution is skewed to the right.
The summary statistics are provided above.

```{r echo=FALSE}
ggplot(wine,aes(total.sulfur.dioxide)) +
  geom_histogram(color='white',binwidth=5) +
  scale_x_continuous(breaks=seq(0,300,25)) +
  xlab('Total Sulfur Dioxide (mg/L)') +
  ylab('Count')

summary(wine$total.sulfur.dioxide)
```

The total sulfur dioxide concentration distribution is skewed to the right.
The summary statistics are provided above.

```{r echo=FALSE}
ggplot(wine,aes(density)) +
  geom_histogram(color='white',binwidth=0.001) +
  scale_x_continuous(breaks=seq(0.99,1.1,0.001)) +
  xlab('Density (g/mL)') +
  ylab('Count')

summary(wine$total.sulfur.dioxide)
```

The density distribution is close to normal distribution. 
The summary statistics are provided above.

```{r echo=FALSE}
ggplot(wine,aes(pH)) +
  geom_histogram(color='white',binwidth=0.04) +
  scale_x_continuous(breaks=seq(2.740,4.010,0.1)) +
  xlab('pH') +
  ylab('Count')

summary(wine$pH)
```

The pH distribution is close to normal distribution.
The summary statistics are provided above.

```{r echo=FALSE}
ggplot(wine,aes(sulphates)) +
  geom_histogram(color='white',binwidth=0.1) +
  scale_x_continuous(breaks=seq(0.3,2.00,0.1)) +
  xlab('Potassium Sulphate (g/L)') +
  ylab('Count')

summary(wine$sulphates)
```

The potassium sulphate concentration distribution is skewed to the right.
The summary statistics are provided above.

```{r echo=FALSE}
ggplot(wine,aes(alcohol)) +
  geom_histogram(color='white',binwidth=0.1) +
  scale_x_continuous(breaks=seq(8.40,14.90,1)) +
  xlab('Alcohol (%)') +
  ylab('Count')

summary(wine$alcohol)
```

The % alcohol distribution is very skewed to the right.
The summary statistics are provided above.

```{r echo=FALSE}
wine$total.acid <- wine$fixed.acidity + wine$volatile.acidity + wine$citric.acid

ggplot(wine,aes(total.acid)) +
  geom_histogram(color='white',binwidth=0.5) +
  scale_x_continuous(breaks=seq(0,22,1)) +
  xlab('Total Acid (g/L)') +
  ylab('Count')

summary(wine$total.acid)
```

Total acid is a new feature calculated from the sum of tartaric acid, 
acetic acid, and citric acid, concentration. The total salt distribution is 
skewed to the right. The summary statistics are provided above.

```{r echo=FALSE}
wine$total.salt <- wine$sulphates + wine$chlorides

ggplot(wine,aes(total.salt)) +
  geom_histogram(color='white',binwidth=0.1) +
  scale_x_continuous(breaks=seq(0.3,3,0.1)) +
  xlab('Total Salt (g/L)') +
  ylab('Count')

summary(wine$total.salt)
```

Total salt is a new feature calculated from the sum of sodium chloride and 
potassium sulphate concentration. The total salt distribution is skewed to the 
right. The summary statistics are provided above.


```{r echo=FALSE}
wine$dioxide.ratio <- wine$free.sulfur.dioxide/wine$total.sulfur.dioxide

ggplot(wine,aes(dioxide.ratio)) +
  #geom_histogram(color='white') +
  geom_histogram(color='white',binwidth=0.02) +
  #scale_x_continuous() +
  scale_x_continuous(breaks=seq(0,1,0.05)) +
  xlab('Sulfur Dioxide Ratio') +
  ylab('Count')

summary(wine$dioxide.ratio)
```

Dioxide ratio is a new feature calculated from dividing free sulfur dioxide by 
total sulfur dioxide concentration. The ratio distribution is close to normal, 
except that there are large number of samples with ratio ~0.23-0.25. 
The summary statistics are provided above.

# Univariate Analysis

### What is the structure of your dataset?
There are 1599 wine samples with 11 features: tartaric acid concentration,
acetic acid concentration, citric acid concentration, residual sugar 
concentration, sodium chloride concentration, free sulfur dioxide concentration,
total sulfur dioxide concentration, density, pH, potassium sulphate 
concentration, % alcohol, and quality. All of the features are continuous 
variables. Some of the feature distributions are close to normal, while others 
are skewed to the right. The wine quality is a score between 0 and 10. 

### What is/are the main feature(s) of interest in your dataset?
The feature of interest is the quality of the wine. This analysis aims to 
investigate which of the other features influence the quality.

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?
All the features are related with the chemical composition, chemical property,
and physical property of the wine sample. So, I suspect each feature will have
some influence on the quality, although some features may have more influence
than the others.

### Did you create any new variables from existing variables in the dataset?
There are 3 new variables created:
1. Total acid concentration: tartaric acid + acetic acid + citric acid 
concentration
2. Total salt concentration: sodium chloride concentration + potassium sulphate
concentration
3. Dioxide ratio: free sulfur dioxide concentration/total sulfur dioxide
concentration

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?
Some of the features are skewed and have 0 values. However, no transformation
was performed as these values are reasonable and will be used as input variables
in regression later.

# Bivariate Plots Section

```{r echo=FALSE}
cor.mtest <- function(mat, conf.level = 0.95){
  mat <- as.matrix(mat)
    n <- ncol(mat)
    p.mat <- lowCI.mat <- uppCI.mat <- matrix(NA, n, n)
    diag(p.mat) <- 0
    diag(lowCI.mat) <- diag(uppCI.mat) <- 1
    for(i in 1:(n-1)){
        for(j in (i+1):n){
            tmp <- cor.test(mat[,i], mat[,j], conf.level = conf.level)
            p.mat[i,j] <- p.mat[j,i] <- tmp$p.value
            lowCI.mat[i,j] <- lowCI.mat[j,i] <- tmp$conf.int[1]
            uppCI.mat[i,j] <- uppCI.mat[j,i] <- tmp$conf.int[2]
        }
    }
    return(list(p.mat, lowCI.mat, uppCI.mat))
}

M = cor(wine[,c(2:16)])
res1 <- cor.mtest(wine[,c(2:16)],0.95)
res2 <- cor.mtest(wine[,c(2:16)],0.99)
## specialized the insignificant value according to the significant level
corrplot(M, p.mat = res1[[1]], sig.level=0.2)

```

The above correlation plot shows that the quality is positively correlated 
with alcohol, potassium sulphate, citric acid, and dioxide ratio. 
Moreover, quality has negative correlation with acetic acid concentration, 
density, and total sulfur dioxide concentration.

Among the features other than quality, the followings are very strong positive
correlations:
1. Tartaric acid and citric acid concentration
2. Tartaric acid concentration and density
3. Free sulfur dioxide and total sulfur dioxide concentration

Similarly, the followings are the strong negative correlations:
1. Citric acid and acetic acid concentration
2. Density and alcohol concentration

Note that the obvious correlations, such as between acid concentration and pH, 
as well as the new features and the features used to derive them, are not 
listed above.

These correlations are analyzed more closely in the following plots.

```{r echo=FALSE}
ggplot(wine,aes(fixed.acidity,citric.acid)) +
  geom_point(alpha=0.5,position='jitter') +
  ylab('Citric Acid (g/L)') +
  xlab('Tartaric Acid (g/L)')

cor.test(wine$citric.acid,wine$fixed.acidity,method='pearson')
```

Both the scatter plot and pearson correlation indicate strong positive 
correlation. It may be interesting to investigate this later when we perform 
multivariate analysis together with quality.

```{r echo=FALSE}
ggplot(wine,aes(fixed.acidity,density)) +
  geom_point(alpha=0.5,position='jitter') +
  ylab('Density (g/mL)') +
  xlab('Tartaric Acid (g/L)')

cor.test(wine$density,wine$fixed.acidity,method='pearson')
```

Both the scatter plot and pearson correlation indicate strong positive 
correlation. We will also revisit this later in multivariate analysis with 
quality.

```{r echo=FALSE}
ggplot(wine,aes(free.sulfur.dioxide,total.sulfur.dioxide)) +
  geom_point(alpha=0.5,position='jitter') +
  ylab('Total Sulfur Dioxide (g/L)') +
  xlab('Free Sulfur Dioxide (g/L)')

cor.test(wine$free.sulfur.dioxide,wine$total.sulfur.dioxide,method='pearson')
```

Similarly, both the scatter plot and pearson correlation indicate strong 
positive correlation. We will also revisit this later in multivariate analysis 
withquality.

```{r echo=FALSE}
ggplot(wine,aes(volatile.acidity,citric.acid)) +
  geom_point(alpha=0.5,position='jitter') +
  ylab('Citric Acid (g/L)') +
  xlab('Acetic Acid (g/L)')

cor.test(wine$volatile.acidity,wine$citric.acid,method='pearson')
```

Both the scatter plot and pearson correlation indicate strong negative
correlation. We will also revisit this later in multivariate analysis with 
quality.

```{r echo=FALSE}
ggplot(wine,aes(alcohol,density)) +
  geom_point(position='jitter',alpha=0.5) +
  ylab('Density (g/mL)') +
  xlab('Alcohol (%)')

cor.test(wine$alcohol,wine$density,method='pearson')
```

The above plot shows that some samples with the same %alcohol seem to 
significantly vary in density. We will also revisit this later in multivariate 
analysis with quality.

The following plots explore the correlation among quality and the other 
features. 

```{r echo=FALSE}
ggplot(wine,aes(quality,fixed.acidity,group=quality)) + 
  geom_boxplot() +
  stat_summary(fun.y = mean,geom='point',shape=4) +
  xlab('Quality') +
  ylab('Tartaric Acid (g/L)')

tapply(wine$fixed.acidity,wine$quality,summary)
```

The boxplot and summary statistics suggest that there is no strong correlation 
between the quality and tartaric acid concentration.

```{r echo=FALSE}
ggplot(wine,aes(quality,volatile.acidity,group=quality)) + 
  geom_boxplot() +
  stat_summary(fun.y = mean,geom='point',shape=4) +
  xlab('Quality') +
  ylab('Acetic Acid (g/L)') 

tapply(wine$volatile.acidity,wine$quality,summary)
```

The boxplot and summary statistics suggest that there may be a strong negative
correlation between the quality and acetic acid concentration.

```{r echo=FALSE}
ggplot(wine,aes(quality,citric.acid,group=quality)) + 
  geom_boxplot() +
  stat_summary(fun.y = mean,geom='point',shape=4) +
  xlab('Quality') +
  ylab('Citric Acid (g/L)') 

tapply(wine$citric.acid,wine$quality,summary)
```

The boxplot and summary statistics suggest that there may be a strong positive
correlation between the quality and citric acid concentration.

```{r echo=FALSE}
ggplot(wine,aes(quality,total.acid,group=quality)) + 
  geom_boxplot() +
  stat_summary(fun.y = mean,geom='point',shape=4) +
  xlab('Quality') +
  ylab('Total Acid (g/L)') 

tapply(wine$total.acid,wine$quality,summary)
```

The boxplot and summary statistics suggest that there is no strong correlation 
between the quality and total acid concentration. This is not surprising since
quality seems to be positively correlated with citric acid concentration, 
negatively correlated with acetic acid concentration, and not correlated with
tartaric acid concentration. So, the total acid concentration may not be 
correlated with the quality.

```{r echo=FALSE}
ggplot(wine,aes(quality,residual.sugar,group=quality)) + 
  geom_boxplot() +
  stat_summary(fun.y = mean,geom='point',shape=4) +
  xlab('Quality') +
  ylab('Residual Sugar (g/L)') 

tapply(wine$residual.sugar,wine$quality,summary)
```

The boxplot and summary statistics suggest that there is no strong correlation 
between the quality and residual sugar concentration. 

```{r echo=FALSE}
ggplot(wine,aes(quality,chlorides,group=quality)) + 
  geom_boxplot() +
  stat_summary(fun.y = mean,geom='point',shape=4) +
  xlab('Quality') +
  ylab('Sodium Chloride (g/L)') 

tapply(wine$chlorides,wine$quality,summary)
```

The boxplot and summary statistics suggest that there is no strong correlation 
between the quality and sodium chloride concentration. 

```{r echo=FALSE}
ggplot(wine,aes(quality,free.sulfur.dioxide,group=quality)) + 
  geom_boxplot() +
  stat_summary(fun.y = mean,geom='point',shape=4) +
  xlab('Quality') +
  ylab('Free Sulfur Dioxide (mg/L)') 

tapply(wine$free.sulfur.dioxide,wine$quality,summary)
```

The boxplot and summary statistics suggest that there is no strong correlation 
between the quality and free sulfur dioxide concentration. 

```{r echo=FALSE}
ggplot(wine,aes(quality,total.sulfur.dioxide,group=quality)) + 
  geom_boxplot() +
  stat_summary(fun.y = mean,geom='point',shape=4) +
  xlab('Quality') +
  ylab('Total Sulfur Dioxide (mg/L)') 

tapply(wine$total.sulfur.dioxide,wine$quality,summary)
```

The boxplot and summary statistics suggest that there is no strong correlation 
between the quality and total sulfur dioxide concentration. 

```{r echo=FALSE}
ggplot(wine,aes(quality,dioxide.ratio,group=quality)) + 
  geom_boxplot() +
  stat_summary(fun.y = mean,geom='point',shape=4) +
  xlab('Quality') +
  ylab('Sulfur Dioxide Ratio') 

tapply(wine$dioxide.ratio,wine$quality,summary)
```

The boxplot and summary statistics suggest that there is no strong correlation 
between the quality and the created sulfur dioxide ratio feature. 

```{r echo=FALSE}
ggplot(wine,aes(quality,density,group=quality)) + 
  geom_boxplot() +
  stat_summary(fun.y = mean,geom='point',shape=4) +
  xlab('Quality') +
  ylab('Density (g/L)') 

tapply(wine$density,wine$quality,summary)
```

The boxplot and summary statistics suggest that there may be some correlation 
between the quality and density - the lower the density, the higher the quality.

```{r echo=FALSE}
ggplot(wine,aes(quality,alcohol,group=quality)) + 
  geom_boxplot() +
  stat_summary(fun.y = mean,geom='point',shape=4) +
  xlab('Quality') +
  ylab('Alcohol (%)') 

tapply(wine$alcohol,wine$quality,summary)
```

The boxplot and summary statistics suggest that there may be some positive 
correlation between the quality and %alcohol.

```{r echo=FALSE}
ggplot(wine,aes(quality,pH,group=quality)) + 
  geom_boxplot() +
  stat_summary(fun.y = mean,geom='point',shape=4) +
  xlab('Quality') +
  ylab('pH') 

tapply(wine$pH,wine$quality,summary)
```

The boxplot and summary statistics suggest that there is no strong correlation 
between the quality and total acid concentration.

```{r echo=FALSE}
ggplot(wine,aes(quality,sulphates,group=quality)) + 
  geom_boxplot() +
  stat_summary(fun.y = mean,geom='point',shape=4) +
  xlab('Quality') +
  ylab('Potassium Sulphate (g/L)') 

tapply(wine$sulphates,wine$quality,summary)
```

The boxplot and summary statistics suggest that there may be some positive 
correlation between the quality and potassium sulphate concentration.

```{r echo=FALSE}
ggplot(wine,aes(quality,total.salt,group=quality)) + 
  geom_boxplot() +
  stat_summary(fun.y = mean,geom='point',shape=4) +
  xlab('Quality') +
  ylab('Total Salt (g/L)') 

tapply(wine$total.salt,wine$quality,summary)
```

The boxplot and summary statistics suggest that there may be some positive 
correlation between the quality and total salt concentration.

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

The quality is positively correlated with alcohol, potassium sulphate, 
citric acid, and dioxide ratio. Moreover, quality has negative correlation with 
acetic acid concentration, density, and total sulfur dioxide concentration.

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

Among the features other than quality, the followings are very strong positive
correlations:
1. Tartaric acid and citric acid concentration
2. Tartaric acid concentration and density
3. Free sulfur dioxide and total sulfur dioxide concentration

Similarly, the followings are the strong negative correlations:
1. Citric acid and acetic acid concentration
2. Density and alcohol concentration

Note that the obvious correlations, such as between acid concentration and pH, 
as well as the new features and the features used to derive them, are not 
listed above.

### What was the strongest relationship you found?

The two correlations between total acid concentration and density as well as
total acid concentration and pH seem to be equally strong. They are the most
obvious correlation observed.

# Multivariate Plots Section

Here, we will re-evaluate the correlations that we found in the bivariate 
analysis by analyzing the correlation between quality and 2 other features.

```{r echo=FALSE, Multivariate_Plots}
ggplot(wine,aes(total.acid,density,color=factor(quality))) +
  geom_point(alpha=0.8,size=1.5) +
  scale_color_brewer(type='div',
                     guide = guide_legend(title='Quality',reverse=FALSE,
                                          override.aes=list(alpha=1,size=2))) +
  ylab('Density (g/mL)') +
  xlab('Total Acid (g/L)') +
  ggtitle('Quality vs Total Acid Concentration and Density') +
  theme_dark()
```

In the above figure, most of the high quality wine samples are on the bottom 
right region, while the low quality on the upper left region.
This plot aligns with the bivariate plot between quality and density. 

```{r echo=FALSE}
ggplot(wine,aes(total.acid,pH,color=factor(quality))) +
  geom_point(alpha=0.8,size=1.5) +
  scale_color_brewer(type='div',
                     guide = guide_legend(title='Quality',reverse=FALSE,
                                          override.aes=list(alpha=1,size=2))) +
  ylab('pH') +
  xlab('Total Acid (g/L)') +
  ggtitle('Quality vs Total Acid Concentration and pH') +
  theme_dark()
```

In the above plot, there is no clear separation between high quality and
low quality wine samples. This observation also aligns with with the 
bivariate analysis, in which the correlation between quality and pH or 
total acid concentration seems to be weak.

```{r echo=FALSE}
ggplot(wine,aes(citric.acid,alcohol,color=factor(quality))) +
  geom_point(alpha=0.8,size=1.5) +
  scale_color_brewer(type='div',
                     guide = guide_legend(title='Quality',reverse=FALSE,
                                          override.aes=list(alpha=1,size=2))) +
  ylab('Alcohol (%)') +
  xlab('Citric Acid (g/L)') +
  ggtitle('Quality vs Citric Acid and Alcohol Concentration') +
  theme_dark()
```

Here, high quality wine samples have high citric acid concentration and
alcohol concentration (upper right region), while low quality samples have low 
citric acid and alcohol concentration (lower left region) aligned with the 
observation in bivariate analysis.

```{r echo=FALSE}
ggplot(wine,aes(volatile.acidity,alcohol,color=factor(quality))) +
  geom_point(alpha=0.8,size=1.5) +
  scale_color_brewer(type='div',
                     guide = guide_legend(title='Quality',reverse=FALSE,
                                          override.aes=list(alpha=1,size=2))) +
  ylab('Alcohol (%)') +
  xlab('Acetic Acid (g/L)') +
  ggtitle('Quality vs Alcohol and Acetic Acid Concentration') +
  theme_dark()
```

Here, most of the high quality wine samples have high alcohol and low acetic 
acid concentration (upper left region), aligned with the observation in 
bivariate analysis.

```{r echo=FALSE}
ggplot(wine,aes(volatile.acidity,density,color=factor(quality))) +
  geom_point(alpha=0.8,size=1.5) +
  scale_color_brewer(type='div',
                     guide = guide_legend(title='Quality',reverse=FALSE,
                                          override.aes=list(alpha=1,size=2))) +
  ylab('Density (g/mL)') +
  xlab('Acetic Acid (g/L)') +
  ggtitle('Quality vs Density and Acetic Acid Concentration') +
  theme_dark()
```

Here, most of the high quality wine samples have low density and acetic acid
concentration (bottom left region), aligned with the observation in bivariate 
analysis.

The quality's positive correlation with citric acid, alcohol, and potassium 
sulphate concentration, as well as negative correlation with acetic acid 
concentration and density shown above suggest that we may be able to build a 
model of quality as a function of these features.

```{r echo=FALSE}
library(memisc)
m1 <- lm(quality ~ alcohol,data=wine)
m2 <- update(m1, ~ . + citric.acid)
m3 <- update(m2, ~ . + sulphates)
m4 <- update(m3, ~ . + volatile.acidity)
m5 <- update(m4, ~ . + density)
mtable(m1,m2,m3,m4,m5,sdigits=3)
```


# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!