Red Wine Quality Exploratory Data Analysis by Lam Raga A. Markely
========================================================


```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(corrplot)
library(psych)
library(ggplot2)
library(dplyr)
library(Amelia)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
wine = read.csv('wineQualityReds.csv')
```

This dataset contains information about various red wine samples with their 
chemical properties and quality. The goal of the analysis is to identify which 
chemical properties influence the quality of red wines.

# Univariate Plots Section

```{r echo=FALSE, Univariate_Plots}
dim(wine)
str(wine)
summary(wine)
```

The red wine dataset consists of 12 features, including 11 input variables 
(fixed.acidity to alcohol), 1 output variable (quality), and 1599 samples. 
The feature X is the index of the dataset. 
The definition of the variables are described in detail 
[here](https://s3.amazonaws.com/udacity-hosted-downloads/ud651/wineQualityInfo.txt).
In the following analysis, the variable fixed acidity will be referred as
tartaric acid, volatile acidity as acetic acid, chlorides as sodium chloride,
and sulphates as potassium sulphate as described in the above link. 

```{r echo=FALSE}
ggplot(wine,aes(quality)) +
  geom_histogram(binwidth=1,color='white') +
  scale_x_continuous(breaks=seq(3,8,1)) +
  xlab('Quality') +
  ylab('Count')

summary(wine$quality)
```

Most of the wine samples have quality 5 and 6. 
The summary statistics are provided above.

```{r echo=FALSE}
ggplot(wine,aes(fixed.acidity)) +
  geom_histogram(color='white',binwidth=1) +
  scale_x_continuous(breaks=seq(4,16,1)) +
  xlab('Tartaric Acid (g/L)') +
  ylab('Count')

summary(wine$fixed.acidity)

```

The tartaric acid concentration distribution is slightly skewed to the right.
The summary statistics are provided above.

```{r echo=FALSE}
ggplot(wine,aes(volatile.acidity)) +
  geom_histogram(color='white',binwidth=0.1) +
  scale_x_continuous(breaks=seq(0.1,1.6,0.1)) +
  xlab('Acetic Acid (g/L)') +
  ylab('Count')

summary(wine$volatile.acidity)
```

The acetic acid concentration distribution is slightly skewed to the right.
The summary statistics are provided above.

```{r echo = FALSE}
ggplot(wine,aes(citric.acid)) +
  geom_histogram(color='white',binwidth=0.05) +
  scale_x_continuous(breaks=seq(0,1.0,0.1)) +
  xlab('Citric Acid (g/L)') +
  ylab('Count')

summary(wine$citric.acid)
```

The citric acid concentration distribution is very skewed to the right.
The summary statistics are provided above.

```{r echo=FALSE}
ggplot(wine,aes(residual.sugar)) +
  geom_histogram(color='white',binwidth=0.5) +
  scale_x_continuous(breaks=seq(0,16,1)) +
  xlab('Residual Sugar (g/L)') +
  ylab('Count')

summary(wine$residual.sugar)
```

The residual sugar concentration distribution is skewed to the right.
The summary statistics are provided above.

```{r echo=FALSE}
ggplot(wine,aes(chlorides)) +
  geom_histogram(color='white',binwidth=0.02) +
  scale_x_continuous(breaks=seq(0,0.7,0.05)) +
  xlab('Sodium Chloride (g/L)') +
  ylab('Count')

summary(wine$chlorides)
```

The sodium chloride concentration distribution is skewed to the right.
The summary statistics are provided above.

```{r echo=FALSE}
ggplot(wine,aes(free.sulfur.dioxide)) +
  geom_histogram(color='white',binwidth=2) +
  scale_x_continuous(breaks=seq(0,80,5)) +
  xlab('Free Sulfur Dioxide (mg/L)') +
  ylab('Count')

summary(wine$free.sulfur.dioxide)
```

The free sulfur dioxide concentration distribution is skewed to the right.
The summary statistics are provided above.

```{r echo=FALSE}
ggplot(wine,aes(total.sulfur.dioxide)) +
  geom_histogram(color='white',binwidth=5) +
  scale_x_continuous(breaks=seq(0,300,25)) +
  xlab('Total Sulfur Dioxide (mg/L)') +
  ylab('Count')

summary(wine$total.sulfur.dioxide)
```

The total sulfur dioxide concentration distribution is skewed to the right. 
The summary statistics are provided above.

```{r echo=FALSE}
ggplot(wine,aes(density)) +
  geom_histogram(color='white',binwidth=0.001) +
  scale_x_continuous(breaks=seq(0.99,1.1,0.001)) +
  xlab('Density (g/mL)') +
  ylab('Count')

summary(wine$total.sulfur.dioxide)
```

The density distribution is close to normal distribution. 
The summary statistics are provided above.

```{r echo=FALSE}
ggplot(wine,aes(pH)) +
  geom_histogram(color='white',binwidth=0.04) +
  scale_x_continuous(breaks=seq(2.740,4.010,0.1)) +
  xlab('pH') +
  ylab('Count')

summary(wine$pH)
```

The pH distribution is close to normal distribution.
The summary statistics are provided above.

```{r echo=FALSE}
ggplot(wine,aes(sulphates)) +
  geom_histogram(color='white',binwidth=0.1) +
  scale_x_continuous(breaks=seq(0.3,2.00,0.1)) +
  xlab('Potassium Sulphate (g/L)') +
  ylab('Count')

summary(wine$sulphates)
```

The potassium sulphate concentration distribution is skewed to the right.
The summary statistics are provided above.

```{r echo=FALSE}
ggplot(wine,aes(alcohol)) +
  geom_histogram(color='white',binwidth=0.1) +
  scale_x_continuous(breaks=seq(8.40,14.90,1)) +
  xlab('Alcohol (%)') +
  ylab('Count')

summary(wine$alcohol)
```

The % alcohol distribution is very skewed to the right.
The summary statistics are provided above.

```{r echo=FALSE}
wine$total.acid <- wine$fixed.acidity + wine$volatile.acidity + wine$citric.acid

ggplot(wine,aes(total.acid)) +
  geom_histogram(color='white',binwidth=0.5) +
  scale_x_continuous(breaks=seq(0,22,1)) +
  xlab('Total Acid (g/L)') +
  ylab('Count')

summary(wine$total.acid)
```

Total acid is a new feature calculated from the sum of tartaric acid, 
acetic acid, and citric acid, concentration. The total salt distribution is 
skewed to the right. The summary statistics are provided above.

```{r echo=FALSE}
wine$total.salt <- wine$sulphates + wine$chlorides

ggplot(wine,aes(total.salt)) +
  geom_histogram(color='white',binwidth=0.1) +
  scale_x_continuous(breaks=seq(0.3,3,0.1)) +
  xlab('Total Salt (g/L)') +
  ylab('Count')

summary(wine$total.salt)
```

Total salt is a new feature calculated from the sum of sodium chloride and 
potassium sulphate concentration. The total salt distribution is skewed to the 
right. The summary statistics are provided above.


```{r echo=FALSE}
wine$dioxide.ratio <- wine$free.sulfur.dioxide/wine$total.sulfur.dioxide

ggplot(wine,aes(dioxide.ratio)) +
  #geom_histogram(color='white') +
  geom_histogram(color='white',binwidth=0.02) +
  #scale_x_continuous() +
  scale_x_continuous(breaks=seq(0,1,0.05)) +
  xlab('Sulfur Dioxide Ratio') +
  ylab('Count')

summary(wine$dioxide.ratio)
```

Dioxide ratio is a new feature calculated from dividing free sulfur dioxide by 
total sulfur dioxide concentration. The ratio distribution is close to normal, 
except that there are large number of samples with ratio ~0.23-0.25. 
The summary statistics are provided above.

# Univariate Analysis

### What is the structure of your dataset?
There are 1599 wine samples with 11 features: tartaric acid concentration,
acetic acid concentration, citric acid concentration, residual sugar 
concentration, sodium chloride concentration, free sulfur dioxide concentration,
total sulfur dioxide concentration, density, pH, potassium sulphate 
concentration, % alcohol, and quality. All of the features except quality are 
continuous variables. Some of the feature distributions are close to normal, 
while others are skewed to the right. 
The wine quality is a score between 0 and 10. 

### What is/are the main feature(s) of interest in your dataset?
The feature of interest is the quality of the wine. This analysis aims to 
investigate which of the other features influence the quality.

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?
All the features are related with the chemical composition, chemical property,
and physical property of the wine sample. So, I suspect many of the features 
will have some influence on the quality, although some features may have more 
influence than the others.

### Did you create any new variables from existing variables in the dataset?
There are 3 new variables created:

1. Total acid concentration: tartaric acid + acetic acid + citric acid 
concentration
2. Total salt concentration: sodium chloride concentration + potassium sulphate
concentration
3. Dioxide ratio: free sulfur dioxide concentration/total sulfur dioxide
concentration

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?
Some of the features are skewed and have 0 values. However, no transformation
was performed as these values are reasonable and will be used as input variables
in linear regression later.

# Bivariate Plots Section


```{r CorrelationPlot, fig.width=8, fig.height=8, echo=FALSE}
cor.mtest <- function(mat, conf.level = 0.95){
  mat <- as.matrix(mat)
    n <- ncol(mat)
    p.mat <- lowCI.mat <- uppCI.mat <- matrix(NA, n, n)
    diag(p.mat) <- 0
    diag(lowCI.mat) <- diag(uppCI.mat) <- 1
    for(i in 1:(n-1)){
        for(j in (i+1):n){
            tmp <- cor.test(mat[,i], mat[,j], conf.level = conf.level)
            p.mat[i,j] <- p.mat[j,i] <- tmp$p.value
            lowCI.mat[i,j] <- lowCI.mat[j,i] <- tmp$conf.int[1]
            uppCI.mat[i,j] <- uppCI.mat[j,i] <- tmp$conf.int[2]
        }
    }
    return(list(p.mat, lowCI.mat, uppCI.mat))
}

M = cor(wine[,c(2:16)])
res1 <- cor.mtest(wine[,c(2:16)],0.95)
res2 <- cor.mtest(wine[,c(2:16)],0.99)
## specialized the insignificant value according to the significant level
corrplot(M, p.mat = res1[[1]], sig.level=0.05)

```

This figure shows correlations between the features in the dataset. The values
correspond to the pearson correlation scores, and crosses indicate that the
absolute value of the pearson correlation is insignificant based on p-value of 
0.05. 

The above correlation plot shows that the quality has a strong positive 
correlation with alcohol and a strong negative correlation with acetic acid.

Among the features other than quality, the followings are very strong positive
correlations:

1. Tartaric acid and citric acid concentration
2. Tartaric acid concentration and density
3. Free sulfur dioxide and total sulfur dioxide concentration

Similarly, the followings are the strong negative correlations:

1. Citric acid and acetic acid concentration
2. Density and alcohol concentration

Note that the obvious correlations, such as between acid concentration and pH, 
as well as the new features and the features used to derive them, are not 
listed above.

These correlations are analyzed more closely in the following plots.

```{r echo=FALSE}
ggplot(wine,aes(fixed.acidity,citric.acid)) +
  geom_point(alpha=0.5,position='jitter') +
  ylab('Citric Acid (g/L)') +
  xlab('Tartaric Acid (g/L)')

cor.test(wine$citric.acid,wine$fixed.acidity,method='pearson')
```

Both the scatter plot and pearson correlation indicate strong positive 
correlation. It may be interesting to investigate this later when we perform 
multivariate analysis together with quality.

```{r echo=FALSE}
ggplot(wine,aes(fixed.acidity,density)) +
  geom_point(alpha=0.5,position='jitter') +
  ylab('Density (g/mL)') +
  xlab('Tartaric Acid (g/L)')

cor.test(wine$density,wine$fixed.acidity,method='pearson')
```

Both the scatter plot and pearson correlation indicate strong positive 
correlation. We will also revisit this later in multivariate analysis with 
quality.

```{r echo=FALSE}
ggplot(wine,aes(free.sulfur.dioxide,total.sulfur.dioxide)) +
  geom_point(alpha=0.5,position='jitter') +
  ylab('Total Sulfur Dioxide (g/L)') +
  xlab('Free Sulfur Dioxide (g/L)')

cor.test(wine$free.sulfur.dioxide,wine$total.sulfur.dioxide,method='pearson')
```

Similarly, both the scatter plot and pearson correlation indicate strong 
positive correlation. We will also revisit this later in multivariate analysis 
withquality.

```{r echo=FALSE}
ggplot(wine,aes(volatile.acidity,citric.acid)) +
  geom_point(alpha=0.5,position='jitter') +
  ylab('Citric Acid (g/L)') +
  xlab('Acetic Acid (g/L)')

cor.test(wine$volatile.acidity,wine$citric.acid,method='pearson')
```

Both the scatter plot and pearson correlation indicate strong negative
correlation. We will also revisit this later in multivariate analysis with 
quality.

```{r echo=FALSE}
ggplot(wine,aes(alcohol,density)) +
  geom_point(position='jitter',alpha=0.5) +
  ylab('Density (g/mL)') +
  xlab('Alcohol (%)')

cor.test(wine$alcohol,wine$density,method='pearson')
```

The above plot shows that some samples with the same %alcohol seem to 
significantly vary in density. We will also revisit this later in multivariate 
analysis with quality.

The following plots explore the correlation among quality and the other 
features. 

```{r echo=FALSE}
ggplot(wine,aes(quality,fixed.acidity,group=quality)) + 
  geom_boxplot() +
  stat_summary(fun.y = mean,geom='point',shape=4) +
  xlab('Quality') +
  ylab('Tartaric Acid (g/L)')

tapply(wine$fixed.acidity,wine$quality,summary)
```

The boxplot and summary statistics suggest that there is no strong correlation 
between the quality and tartaric acid concentration.

```{r echo=FALSE}
ggplot(wine,aes(quality,volatile.acidity,group=quality)) + 
  geom_boxplot() +
  stat_summary(fun.y = mean,geom='point',shape=4) +
  xlab('Quality') +
  ylab('Acetic Acid (g/L)') 

tapply(wine$volatile.acidity,wine$quality,summary)
```

The boxplot and summary statistics show a negative
correlation between the quality and acetic acid concentration.

```{r echo=FALSE}
ggplot(wine,aes(quality,citric.acid,group=quality)) + 
  geom_boxplot() +
  stat_summary(fun.y = mean,geom='point',shape=4) +
  xlab('Quality') +
  ylab('Citric Acid (g/L)') 

tapply(wine$citric.acid,wine$quality,summary)
```

The boxplot and summary statistics shows a weak positive
correlation between the quality and citric acid concentration.

```{r echo=FALSE}
ggplot(wine,aes(quality,total.acid,group=quality)) + 
  geom_boxplot() +
  stat_summary(fun.y = mean,geom='point',shape=4) +
  xlab('Quality') +
  ylab('Total Acid (g/L)') 

tapply(wine$total.acid,wine$quality,summary)
```

The boxplot and summary statistics suggest that there is no strong correlation 
between the quality and total acid concentration. This is not surprising since
quality seems to be positively correlated with citric acid concentration, 
negatively correlated with acetic acid concentration, and not correlated with
tartaric acid concentration. So, the total acid concentration may not be 
correlated with the quality.

```{r echo=FALSE}
ggplot(wine,aes(quality,residual.sugar,group=quality)) + 
  geom_boxplot() +
  stat_summary(fun.y = mean,geom='point',shape=4) +
  xlab('Quality') +
  ylab('Residual Sugar (g/L)') 

tapply(wine$residual.sugar,wine$quality,summary)
```

The boxplot and summary statistics suggest that there is no strong correlation 
between the quality and residual sugar concentration. 

```{r echo=FALSE}
ggplot(wine,aes(quality,chlorides,group=quality)) + 
  geom_boxplot() +
  stat_summary(fun.y = mean,geom='point',shape=4) +
  xlab('Quality') +
  ylab('Sodium Chloride (g/L)') 

tapply(wine$chlorides,wine$quality,summary)
```

The boxplot and summary statistics suggest that there is no strong correlation 
between the quality and sodium chloride concentration. 

```{r echo=FALSE}
ggplot(wine,aes(quality,free.sulfur.dioxide,group=quality)) + 
  geom_boxplot() +
  stat_summary(fun.y = mean,geom='point',shape=4) +
  xlab('Quality') +
  ylab('Free Sulfur Dioxide (mg/L)') 

tapply(wine$free.sulfur.dioxide,wine$quality,summary)
```

The boxplot and summary statistics suggest that there is no strong correlation 
between the quality and free sulfur dioxide concentration. 

```{r echo=FALSE}
ggplot(wine,aes(quality,total.sulfur.dioxide,group=quality)) + 
  geom_boxplot() +
  stat_summary(fun.y = mean,geom='point',shape=4) +
  xlab('Quality') +
  ylab('Total Sulfur Dioxide (mg/L)') 

tapply(wine$total.sulfur.dioxide,wine$quality,summary)
```

The boxplot and summary statistics suggest that there is no strong correlation 
between the quality and total sulfur dioxide concentration. 

```{r echo=FALSE}
ggplot(wine,aes(quality,dioxide.ratio,group=quality)) + 
  geom_boxplot() +
  stat_summary(fun.y = mean,geom='point',shape=4) +
  xlab('Quality') +
  ylab('Sulfur Dioxide Ratio') 

tapply(wine$dioxide.ratio,wine$quality,summary)
```

The boxplot and summary statistics suggest that there is no strong correlation 
between the quality and the created sulfur dioxide ratio feature. 

```{r echo=FALSE}
ggplot(wine,aes(quality,density,group=quality)) + 
  geom_boxplot() +
  stat_summary(fun.y = mean,geom='point',shape=4) +
  xlab('Quality') +
  ylab('Density (g/L)') 

tapply(wine$density,wine$quality,summary)
```

The boxplot and summary statistics show a weak negative correlation between 
quality and density.

```{r echo=FALSE}
ggplot(wine,aes(quality,alcohol,group=quality)) + 
  geom_boxplot() +
  stat_summary(fun.y = mean,geom='point',shape=4) +
  xlab('Quality') +
  ylab('Alcohol (%)') 

tapply(wine$alcohol,wine$quality,summary)
```

The boxplot and summary statistics show a positive 
correlation between the quality and %alcohol.

```{r echo=FALSE}
ggplot(wine,aes(quality,pH,group=quality)) + 
  geom_boxplot() +
  stat_summary(fun.y = mean,geom='point',shape=4) +
  xlab('Quality') +
  ylab('pH') 

tapply(wine$pH,wine$quality,summary)
```

The boxplot and summary statistics suggest that there is no strong correlation 
between the quality and total acid concentration.

```{r echo=FALSE}
ggplot(wine,aes(quality,sulphates,group=quality)) + 
  geom_boxplot() +
  stat_summary(fun.y = mean,geom='point',shape=4) +
  xlab('Quality') +
  ylab('Potassium Sulphate (g/L)') 

tapply(wine$sulphates,wine$quality,summary)
```

The boxplot and summary statistics show a positive 
correlation between the quality and potassium sulphate concentration.

```{r echo=FALSE}
ggplot(wine,aes(quality,total.salt,group=quality)) + 
  geom_boxplot() +
  stat_summary(fun.y = mean,geom='point',shape=4) +
  xlab('Quality') +
  ylab('Total Salt (g/L)') 

tapply(wine$total.salt,wine$quality,summary)
```

The boxplot and summary statistics show a weak positive 
correlation between the quality and total salt concentration.

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

The quality has strong positive correlation with alcohol and strong negative 
correlation with acetic acid.

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

Among the features other than quality, the followings are very strong positive
correlations:

1. Tartaric acid and citric acid concentration
2. Tartaric acid concentration and density
3. Free sulfur dioxide and total sulfur dioxide concentration

Similarly, the followings are the strong negative correlations:

1. Citric acid and acetic acid concentration
2. Density and alcohol concentration

Note that the obvious correlations, such as between acid concentration and pH, 
as well as the new features and the features used to derive them, are not 
listed above.

### What was the strongest relationship you found?

The strongest correlation was tartaric acid concentration and density.

# Multivariate Plots Section

Here, we will re-evaluate the correlations that we found in the bivariate 
analysis by analyzing the correlation between quality and 2 other features.

```{r echo=FALSE}
ggplot(wine,aes(fixed.acidity,citric.acid,color=factor(quality))) +
  geom_point(alpha=0.8,size=1.5,position='jitter') +
  scale_color_brewer(type='div',
                     guide = guide_legend(title='Quality',reverse=FALSE,
                                          override.aes=list(alpha=1,size=2))) +
  ylab('Citric Acid (g/L)') +
  xlab('Tartaric Acid (g/L)') +
  ggtitle('Quality vs Tartaric Acid and Citric Acid Concentration') +
  theme_dark()
```

In the above figure, there is no clear separation between high quality and low
quality wine samples. This aligns with the bivariate analysis result in which 
quality did not show strong correlation with tartaric acid and citric acid 
concentration.

```{r echo=FALSE, Multivariate_Plots}
ggplot(wine,aes(fixed.acidity,density,color=factor(quality))) +
  geom_point(alpha=0.8,size=1.5,position='jitter') +
  scale_color_brewer(type='div',
                     guide = guide_legend(title='Quality',reverse=FALSE,
                                          override.aes=list(alpha=1,size=2))) +
  ylab('Density (g/mL)') +
  xlab('Tartaric Acid (g/L)') +
  ggtitle('Quality vs Tartaric Acid Concentration and Density') +
  theme_dark()
```

In the above figure, most of the high quality wine samples are on the bottom 
right region, while the low quality on the upper left region. However, there is
no separation if one analyzes only from one of the axes, which aligns with the
bivariate analysis result in which quality did not show significant correlation
with tartaric acid concentration and density.

```{r echo=FALSE}
ggplot(wine,aes(free.sulfur.dioxide,total.sulfur.dioxide,
                color=factor(quality))) +
  geom_point(alpha=0.8,size=1.5,position='jitter') +
  scale_color_brewer(type='div',
                     guide = guide_legend(title='Quality',reverse=FALSE,
                                          override.aes=list(alpha=1,size=2))) +
  ylab('Total Sulfur Dioxide (g/L)') +
  xlab('Free Sulfur Dioxide (g/L)') +
  ggtitle('Quality vs Free Sulfur Dioxide and Total Sulfur Dioxide') +
  theme_dark()
```

In the above figure, most of the high quality wine samples are on the bottom 
right region, while the low quality on the upper left region. However, there is
no separation if one analyzes only from one of the axes, which aligns with the
bivariate analysis result in which quality did not show strong correlation
with total and free sulfur dioxide concentration. 

```{r echo=FALSE}
ggplot(wine,aes(volatile.acidity,citric.acid,color=factor(quality))) +
  geom_point(alpha=0.8,size=1.5) +
  scale_color_brewer(type='div',
                     guide = guide_legend(title='Quality',reverse=FALSE,
                                          override.aes=list(alpha=1,size=2))) +
  ylab('Citric Acid (g/L)') +
  xlab('Acetic Acid (g/L)') +
  ggtitle('Quality vs Acetic Acid and Citric Acid Concentration') +
  theme_dark()
```

In the above plot, most of the high quality wine samples are on the left, while
the low quality samples are on the right, which aligns with the bivariate 
analysis result in which quality shows negative correlation with acetic acid
concentration.

```{r echo=FALSE}
ggplot(wine,aes(alcohol,density,color=factor(quality))) +
  geom_point(alpha=0.8,size=1.5,position='jitter') +
  scale_color_brewer(type='div',
                     guide = guide_legend(title='Quality',reverse=FALSE,
                                          override.aes=list(alpha=1,size=2))) +
  ylab('Density (g/mL)') +
  xlab('Alcohol (%)') +
  ggtitle('Quality vs Density and Alcohol Concentration') +
  theme_dark()
```

In the above plot, most of the high quality wine samples are on the right, while
the low quality samples are on the left, which aligns with the bivariate 
analysis result in which quality shows positive correlation with alcohol
concentration.

```{r echo=FALSE}
ggplot(wine,aes(volatile.acidity,alcohol,color=factor(quality))) +
  geom_point(alpha=0.8,size=1.5) +
  scale_color_brewer(type='div',
                     guide = guide_legend(title='Quality',reverse=FALSE,
                                          override.aes=list(alpha=1,size=2))) +
  ylab('Alcohol (%)') +
  xlab('Acetic Acid (g/L)') +
  ggtitle('Quality vs Alcohol and Acetic Acid Concentration') +
  theme_dark()
```

Here, most of the high quality wine samples have high alcohol and low acetic 
acid concentration (upper left region). This result aligns with the bivariate
analysis in which quality positively correlates with alcohol and negatively 
correlates with acetic acid.

```{r echo=FALSE}
ggplot(wine,aes(volatile.acidity,sulphates,color=factor(quality))) +
  geom_point(alpha=0.8,size=1.5) +
  scale_color_brewer(type='div',
                     guide = guide_legend(title='Quality',reverse=FALSE,
                                          override.aes=list(alpha=1,size=2))) +
  ylab('Potassium Sulphate (g/L)') +
  xlab('Acetic Acid (g/L)') +
  ggtitle('Quality vs Acetic Acid and Potassium Sulphate Concentration') +
  theme_dark()
```

Here, most of the high quality wine samples are on the upper left region, and 
the low quality are in the lower right region. This result aligns with the 
bivariate analysis in which quality positively correlates with potassium 
sulphate and negatively correlates with acetic acid concentration.

```{r echo=FALSE}
ggplot(wine,aes(alcohol,chlorides,color=factor(quality))) +
  geom_point(alpha=0.8,size=1.5) +
  scale_color_brewer(type='div',
                     guide = guide_legend(title='Quality',reverse=FALSE,
                                          override.aes=list(alpha=1,size=2))) +
  ylab('Sodium Chloride (g/L)') +
  xlab('Alcohol (%)') +
  ggtitle('Quality vs Alcohol and Sodium Chloride Concentration') +
  theme_dark()
```

Here, most of the high quality wine samples are on the right, while the low 
quality samples are on the left. This result aligns with the 
bivariate analysis in which quality positively correlates with alcohol 
concentration.

```{r echo=FALSE}
ggplot(wine,aes(alcohol,pH,color=factor(quality))) +
  geom_point(alpha=0.8,size=1.5) +
  scale_color_brewer(type='div',
                     guide = guide_legend(title='Quality',reverse=FALSE,
                                          override.aes=list(alpha=1,size=2))) +
  ylab('pH') +
  xlab('Alcohol (%)') +
  ggtitle('Quality vs Alcohol Concentration and pH') +
  theme_dark()
```

Here, most of the high quality wine samples are on the right, while the low 
quality samples are on the left. This result aligns with the 
bivariate analysis in which quality positively correlates with alcohol 
concentration.

In the followings, multinomial logistic regression is used to build a predictive
model based on the wine sample features. 

```{r echo=FALSE}
sapply(wine,function(x) sum(is.na(x)))
```
First, confirm that there is no missing value.

Next, we need to use stratified splitting as there is imbalance in the number
of wine samples in each quality category.

The above results show that only the following features are significant:
alcohol, acetic acid, sodium chloride, potassium sulphate, and pH. The following
is the fitting with only these features.


**Insert comment here**

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

Using acetic acid and alcohol concentration in combination to distinguish high 
quality wine from low quality wine is easier than using one of these two 
features. In the bivariate box plots, quality shows a positive correlation with 
acetic acid and alcohol concentration. However, there are overlaps and outliers
that make it difficult to assess the strength of the correlation. 
Combining quality, acetic acid, and alcohol in one multivariate plot shows a 
much better separation between high quality and low quality wine samples.

Similar trend is also observed when analyzing the following multivariate plots:

1. Quality vs alcohol and potassium sulphate concentration
2. Quality vs tartaric acid concentration and density
3. Quality vs citric acid and alcohol concentration
4. Quality vs sodium chloride and alcohol concentration
5. Quality vs pH and alcohol concentration

### Were there any interesting or surprising interactions between features?
Initially, I was surprised that the density has a strong positive correlation 
with tartaric acid concentration. However, I found out that tartaric acid 
density is 1.78 g/mL, much higher than density of wine ~ 1g/L.

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

Yes, a linear model was created with alcohol, acetic acid, sodium chloride, pH,
and potassium sulphate as the input variables. The variables in the model 
account for 34.6% of the variance in the wine sample quality. Alcohol and acetic
acid have the strongest influence as indicated by significant increase in 
adjusted R-square for these variables compared with the other variables. 

------

# Final Plots and Summary

### Plot One
```{r CorrelationPlot, fig.width=8, fig.height=8, echo=FALSE}
cor.mtest <- function(mat, conf.level = 0.95){
  mat <- as.matrix(mat)
    n <- ncol(mat)
    p.mat <- lowCI.mat <- uppCI.mat <- matrix(NA, n, n)
    diag(p.mat) <- 0
    diag(lowCI.mat) <- diag(uppCI.mat) <- 1
    for(i in 1:(n-1)){
        for(j in (i+1):n){
            tmp <- cor.test(mat[,i], mat[,j], conf.level = conf.level)
            p.mat[i,j] <- p.mat[j,i] <- tmp$p.value
            lowCI.mat[i,j] <- lowCI.mat[j,i] <- tmp$conf.int[1]
            uppCI.mat[i,j] <- uppCI.mat[j,i] <- tmp$conf.int[2]
        }
    }
    return(list(p.mat, lowCI.mat, uppCI.mat))
}

M = cor(wine[,c(2:16)])
res1 <- cor.mtest(wine[,c(2:16)],0.95)
res2 <- cor.mtest(wine[,c(2:16)],0.99)
## specialized the insignificant value according to the significant level
corrplot(M, p.mat = res1[[1]], sig.level=0.05)

```

### Description One

Various features affect the quality of wine. Among them, alcohol and acetic acid
concentration have the strongest correlation with the quality.

### Plot Two
```{r echo=FALSE}
ggplot(wine,aes(volatile.acidity,alcohol,color=factor(quality))) +
  geom_point(alpha=0.8,size=1.5) +
  scale_color_brewer(type='div',
                     guide = guide_legend(title='Quality',reverse=FALSE,
                                          override.aes=list(alpha=1,size=2))) +
  ylab('Alcohol (%)') +
  xlab('Acetic Acid (g/L)') +
  ggtitle('Quality vs Alcohol and Acetic Acid Concentration') +
  theme_dark()
```

### Description Two
Most of the high quality wine samples have high alcohol and low acetic 
acid concentration (upper left region), while most of the low quality wine 
samples have low alcohol and high acetic acid concentration (lower right 
region). This correlation makes sense as wine with high acetic acid 
concentration may have vinegar taste, and wine with low alcohol concentration
may not taste good. 

### Plot Three
```{r echo=FALSE}
ggplot(wine,aes(volatile.acidity,sulphates,color=factor(quality))) +
  geom_point(alpha=0.8,size=1.5) +
  scale_color_brewer(type='div',
                     guide = guide_legend(title='Quality',reverse=FALSE,
                                          override.aes=list(alpha=1,size=2))) +
  ylab('Potassium Sulphate (g/L)') +
  xlab('Acetic Acid (g/L)') +
  ggtitle('Quality vs Acetic Acid and Potassium Sulphate Concentration') +
  theme_dark()
```

### Description Three
The next important feature after alcohol and acetic acid concentration is 
potassium phosphate concentration. Using both potassium sulphate and acetic 
acid concentration, we can also distinguish high quality wine (upper left) from
low quality wine (lower right) although this plot has less clear separation
than the second plot.

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

The wine dataset consists of 1599 wine samples with various quality and feature
values. This projects gradually explores the correlation among the features
starting from univariate, bivariate, to multivariate analysis. This analysis
identifies that alcohol and acetic acid concentration are the features that 
have the most significant influence on the wine quality. This conclusion makes 
sense as wine with high acetic acid concentration may have vinegar taste, and 
wine with low alcohol concentration may not taste good. 

The challenges faced in the analysis were the relatively large number of 
outliers observed during bivariate and multivariate analysis. Larger number of
data may help clarify if the outliers are indeed outliers due to some 
measurement errors, or they are in fact the results of more complex 
interactions between the features that cannot be captured by 
simply plotting 2 to 3 variables in a single plot. It may also bring more 
visual clarity in the separation between high quality and low quality samples 
when assessing the results. 

